Leadership across cultures and contexts: an exploratory analysis

### 2020-05-11 leadership_across_cultures_contexts.Rmd
Some inline percents of benefits and costs

The frequency of leader and follower costs and benefits were estimated similarly to the leader qualities and functions described above. In these data leaders most commonly received benefits in the form of greater access to material resources (documented in about 70% of cultures) and increased social status (documented in about 65% of cultures). Mating benefits and social services benefits were also relatively frequent benefits of leadership. In general, the costs tracked the benefits. Leaders were often described as losing their high social status (documented in about 45% of cultures) and greater access to resources (documented in about 40% of cultures). See Figure \@ref(fig:varsupportcb).

Followers on the other hand most commonly benefited from leadership systems through receiving material resources (documented in about 50% of cultures) but otherwise most frequently through receiving social services (documented in about 35% of cultures). Mating and fitness benefits were relatively rare for followers. Followers were most frequently described as incurring costs in the form of material resources lost (documented in about 40% of cultures) and in providing social services (documented in about 25% of cultures). See Figure \@ref(fig:varsupportcb).

@@

Leadership can evolve when the benefits exceed the costs, and the same for followership. Our data did not allow us to compare benefits and costs. Instead, we compared levels of evidence for benefits vs. costs for each leader benefit/cost dimension and each follower benefit/cost dimension. Specifically, we fit a mixed effects logistic regression model of evidence (0/1) as a function of each dimension (e.g., Mating, Food), a term indicating if the evidence was for a benefit or cost, and their interaction, with random intercepts for author nested within culture. We fit separate models for leaders and followers. These two models allowed us to estimate the relative evidence for, e.g., leader mating benefits vs. leader mating costs, and follower mating benefits vs. follower mating costs, etc. For leaders, the odds ratio that a text record would provide evidence for a benefit vs. a cost (averaging across all benefit/cost dimensions), was $OR=`r signif(leader_bc_ratio$benefit_cost_OR, 3)`$, and for followers, was $OR=`r signif(follower_bc_ratio$benefit_cost_OR, 3)`$. Hence for both leaders and followers, there was more evidence of benefits than costs.

Within benefit/cost dimensions, there was relatively more evidence for mating and territory benefits than costs for both leaders and followers, and equal levels of evidence for risk of harm. For leaders compared to followers, there were noticably higher odds ratios for inclusive fitness benefits, material resources, and social services. See Figure \@ref(fig:benefitcostratio).

@@

All evolutionary models of leadership propose that the fitness benefits of leading outweigh the costs. The relative abundance of evidence for leader benefits vs. costs ($OR=`r signif(leader_bc_ratio$benefit_cost_OR, 3)`$), while not providing evidence of net fitness benefits, is nevertheless encouraging, especially since leaders and followers both had more evidence of mating and territory benefits than costs, which are key contributors to fitness, and leaders also had more evidence for inclusive fitness, material and social benefits, relative to costs. See Figure \@ref(fig:benefitcostratio).

>>>>

```{r cultbc}
leader_benefit_material <- signif(100*leader_benefits_cult$Estimate[leader_benefits_cult$Variable == 'Material resources'], 3)

leader_benefit_status <- signif(100*leader_benefits_cult$Estimate[leader_benefits_cult$Variable == 'Increased social status'], 3)

leader_cost_material <- signif(100*leader_costs_cult$Estimate[leader_costs_cult$Variable == 'Loss of material resources'], 3)

leader_cost_status <- signif(100*leader_costs_cult$Estimate[leader_costs_cult$Variable == 'Reduced social status'], 3)

follower_benefit_material <- signif(100*follower_benefits_cult$Estimate[follower_benefits_cult$Variable == 'Material resources'], 3)

follower_benefit_social <- signif(100*follower_benefits_cult$Estimate[follower_benefits_cult$Variable == 'Social services'], 3)

follower_cost_material <- signif(100*follower_costs_cult$Estimate[follower_costs_cult$Variable == 'Loss of material resources'], 3)

follower_cost_social <- signif(100*follower_costs_cult$Estimate[follower_costs_cult$Variable == 'Loss of social services'], 3)
```

The frequency of leader and follower costs and benefits were estimated similarly to the leader qualities and functions described above. In these data leaders most commonly received benefits in the form of greater access to material resources (documented in `r leader_benefit_material`% of cultures) and increased social status (documented in `r leader_benefit_status`% of cultures). Mating benefits and social services benefits were also relatively frequent benefits of leadership. In general, the costs tracked the benefits. Leaders were often described as losing their high social status (documented in `r leader_cost_status`% of cultures) and greater access to resources (documented in `r leader_cost_material`% of cultures). See Figure \@ref(fig:varsupportcb).

Followers on the other hand most commonly benefited from leadership systems through receiving material resources (documented in `r follower_benefit_material`% of cultures) but otherwise most frequently through receiving social services (documented in `r follower_benefit_social`% of cultures). Mating and fitness benefits were relatively rare for followers. Followers were most frequently described as incurring costs in the form of material resources lost (documented in `r follower_cost_material`% of cultures) and in providing social services (documented in `r follower_cost_social`% of cultures). See Figure \@ref(fig:varsupportcb).

@@

Leadership can evolve when the benefits exceed the costs, and the same for followership. Our data did not allow us to compare benefits and costs. Instead, we compared levels of evidence for benefits vs. costs for each leader benefit/cost dimension and each follower benefit/cost dimension. Specifically, we fit a mixed effects logistic regression model of evidence (0/1) as a function of each dimension (e.g., Mating, Food), a term indicating if the evidence was for a benefit or cost, and their interaction, with random intercepts for author nested within culture. We fit separate models for leaders and followers. These two models allowed us to estimate the relative evidence for, e.g., leader mating benefits vs. leader mating costs, and follower mating benefits vs. follower mating costs, etc. For leaders, the odds ratio that a text record would provide evidence for a benefit vs. a cost (averaging across all benefit/cost dimensions), was $OR=`r signif(leader_bc_ratio$benefit_cost_OR, 2)`$, and for followers, was $OR=`r signif(follower_bc_ratio$benefit_cost_OR, 2)`$. Hence for both leaders and followers, there was more evidence of benefits than costs.

Within benefit/cost dimensions, there was relatively more evidence for mating and territory benefits than costs for both leaders and followers, and equal levels of evidence for protection from vs. risk of harm. For leaders compared to followers, there were noticably higher odds ratios for inclusive fitness benefits, material resources, and social services. See Figure \@ref(fig:benefitcostratio).

@@

All evolutionary models of leadership propose that the fitness benefits of leading outweigh the costs. The relative abundance of evidence for leader and follower benefits vs. costs ($OR=`r signif(leader_bc_ratio$benefit_cost_OR, 2)`$ and $OR=`r signif(follower_bc_ratio$benefit_cost_OR, 2)`$, resp.), while not providing evidence of net fitness benefits, is nevertheless encouraging, especially since leaders and followers both had more evidence of mating and territory benefits than costs, which are key contributors to fitness, and leaders also had more evidence for inclusive fitness, material and social benefits, relative to costs. See Figure \@ref(fig:benefitcostratio).
